---
title: "2009 data expo project 2022"
author: '190527129'
date: "3/23/2022"
output: html_document
---

# Set Working Directory
```{r}
setwd("C:/Users/limzy/Documents/R/dataverse_files")
```
#Question 1

# Read in the data for 2007

```{r}
Y2007 <- read.csv("2007.csv")

str(Y2007)

summary(Y2007)

library(tidyverse)
library(lubridate)
```

# Convert the Y2007 dates to year/month/date format 

```{r}
datesof2007 <- Y2007 %>% select(Year, Month,DayofMonth)
datesof2007 <- mutate(datesof2007,Dates=paste(Year,Month,DayofMonth,sep="/"))
datesof2007 = subset(datesof2007, select = -c(Year, Month, DayofMonth))
datesof2007 = ymd(datesof2007$Dates)
```

# Adding Y2007 dates & day of the week to flight.timing df

```{r}
flight.timingsof2007 <- Y2007 %>% select(ArrDelay)
flight.timingsof2007$datesof2007 <- datesof2007
weekday <- weekdays(datesof2007)
flight.timingsof2007$weekday <- weekday
```

# Reduce the size of the data by finding the mean of the delays per week

```{r}
weeksof2007 <- strftime( datesof2007, format = "%V")    # Converts the time to character objects
weeksof2007 <- as.numeric(weeksof2007)
flight.timingsof2007 <- cbind(flight.timingsof2007, weeksof2007)
```

# Preparation of data for loop function to find the mean delay of the 52 weeks

```{r}
WeeklyArrDelayMean <- vector(mode = "integer", length = length(1:52)) # Empty vector to store values for loop
```

# Loop for mean delay of the 52 weeks

```{r}
for (i in 1:52) {
  WeekArrDelayMeandata <- flight.timingsof2007 %>% filter(weeksof2007 == i)
  
  WeeklyArrDelayMean[i] <- mean(WeekArrDelayMeandata$ArrDelay, na.rm = TRUE)
  
} 
```

# Preparing necessary vectors for plot

```{r}
Numofweeks <- 1:52 # Generating number of weeks vector

weeklyMeanDelaysof2007 <- cbind(WeeklyArrDelayMean, Numofweeks )

weeklyMeanDelaysof2007 <- as.data.frame(weeklyMeanDelaysof2007)
```

# Plotting for mean delay of the 52 weeks

```{r}
w <- ggplot(weeklyMeanDelaysof2007,
            aes(x = Numofweeks,
                y = WeeklyArrDelayMean)
                         )
```

# Line plot

```{r}
w <- w + 
  geom_line()
```

# Plot Labels and plot the graph

```{r}
w <- w + 
  labs(
    title = "Weekly Delays for 2007",
    x = "Number of weeks",
    y = "Mean of Weekly Delay"
  )

w

```

# Preparation of data for loop function to find daily mean delays

# Generating data frame

```{r}
days <- yday(datesof2007) - 1 # Generating numbered days of the year from dates

days_of_delay <- Y2007 %>% select(ArrDelay) %>% cbind(days) 

dailyArrDelayMean <- vector(mode = "integer", length = length(0:364)) # Empty vector to store values for loop
```

# Loop for daily mean of delays

```{r}
for (i in 0:364) {
  dailyArrDelayMeandata <- days_of_delay %>% filter(days == i)
  
  dailyArrDelayMean[i] <- mean(dailyArrDelayMeandata$ArrDelay, na.rm = TRUE) 
  
} 
```

# Preparing necessary vectors for plot

```{r}
ymd_of_2007 <- seq(as.Date("2007-01-01"),as.Date("2007-12-31"),by="1 day") # Generating Year 2007 dates

day_of_the_week <- wday(ymd_of_2007, label = TRUE) # Days

months_of_2007 <- month(ymd_of_2007, label = TRUE) # Months
```

# Combining vectors into data frame without compromising class

```{r}
dailyArrDelayMean <- data.frame(dailyArrDelayMean = dailyArrDelayMean,
                                ymd_of_2007 = ymd_of_2007,
                                months_of_2007 = months_of_2007,
                                day_of_the_week = day_of_the_week)

```

# Plotting daily mean delays 2007

```{r}
p <- ggplot(data = dailyArrDelayMean,
            mapping = aes(x =months_of_2007,
                          y =day_of_the_week,
                          fill =dailyArrDelayMean))
```

# heatmap

```{r}
p <- p + 
  geom_tile()
```

# Colours for heatmap   

```{r}
p <- p + 
  scale_fill_distiller(limits = c(-40, 40),
                       palette = 'RdBu')
```

# Labels for heatmap and plot

```{r}
p <- p +
  labs(
    title = "Daily Delays for 2007",
    fill = "Intensity of \n mean delays",
    x = "Month",
    y = "Day"
  )

p
```

# Best time of the day
# Lowest mean delays from heat map is
# October - Monday

```{r}
CRSDepTime2007 <- Y2007$CRSDepTime
Month_2007 <- Y2007$Month
Delays_2007 <- Y2007$ArrDelay


Scheduled_DepTiming2007 <- data.frame(CRSDepTime2007 = CRSDepTime2007,
                                      Delays_2007 = Delays_2007,
                                      Month_2007 = Month_2007,
                                      weekday = weekday)
``` 

# Filter flights on October Mondays out

```{r}
OctMonScheduled_DepTiming2007 <- filter(Scheduled_DepTiming2007, Month_2007 == 10)

OctMonScheduled_DepTiming2007 <- filter(OctMonScheduled_DepTiming2007, weekday == "Monday")

```

# Scatter plot

```{r}
q <- ggplot(OctMonScheduled_DepTiming2007,
            aes(x = CRSDepTime2007,
                y = Delays_2007)
)
q <- q + 
  geom_point()
```

# Labels for scatter plot

```{r}
q <- q +
  labs(
    title = "October, Monday departures for 2007",
    x = "Time (Hrs)",
    y = "Delays (in Mins)"
  )
```

# Scale for plot and plot

```{r}
q <- q + 
  scale_x_continuous(limits = c(0,2400), breaks = 200*(0:12)) +
  scale_y_continuous(limits = c(-80,200), breaks = 20*(-10:10))

q
```

############
# Question 2

# Read in data

```{r}
planedata <- read.csv("plane-data.csv")

planedelays <- Y2007 %>% select(TailNum, ArrDelay)
```

# Remove rows with NA values in specific column

```{r}
planedelays <- planedelays %>% filter_at(vars(ArrDelay), all_vars(!is.na(.)))
```

# Group by tail number and find the mean of the delays for each tail number

```{r}
mean_plane_delays <- planedelays %>% group_by(TailNum) %>% summarise(mean_delays = mean(ArrDelay))
```

# Merging 2 data frames together by their tail number

```{r}
merged_tailnum_delays <- merge(mean_plane_delays, planedata, by.x = "TailNum", by.y = "tailnum")                                                                 

merged_tailnum_delays <- merged_tailnum_delays %>% select(TailNum, mean_delays, year)
```

# Plotting daily mean delays 2007

```{r}
r <- ggplot(data = merged_tailnum_delays,
            mapping = aes(x =mean_delays,
                          y =year
                          ))
```

# Scatter plot

```{r}
r <- r + 
  geom_point()
```

# Labels for scatter plot

```{r}
r <- r +
  labs(
    title = "2007 Delays of against age of aircrafts",
    x = "Mean delays (mins)",
    y = "Year of aircraft issued"
  )
```

# Scale for plot and plot

```{r}
r <- r + 
  scale_y_discrete(limits = factor(1956:2007), breaks = 2*(978:1004)) +
  scale_x_continuous(limits = c(-50,100), breaks = 10*(-5:100))


r
```

# old planes < 1996

```{r}
OLDmerged_tailnum_delays <- merged_tailnum_delays %>% filter(year < 1996)

OLDmerged_tailnum_delays %>% summarise(mean_delays = mean(mean_delays)) #12.93
```

#new planes =>1996

```{r}
NEWmerged_tailnum_delays <- merged_tailnum_delays %>% filter(year > 1995)

NEWmerged_tailnum_delays %>% summarise(mean_delays = mean(mean_delays)) #10.71
```

############

# Question 3

```{r}
airportdata <- read.csv("airports.csv")

location_of_flights <- Y2007 %>% select(Origin, Dest, FlightNum, UniqueCarrier)

location_of_flights <- merge(airportdata, location_of_flights, by.x = 'iata', by.y = 'Origin')
```

# Number of flights per airport origin

```{r}
Num_of_flights_per_airport <- location_of_flights %>%
  group_by(airport) %>%
  summarize(num_flights = n()) 
```

# Merge data frames

```{r}
Num_of_flights_per_airport <- merge(Num_of_flights_per_airport, airportdata, by.x = 'airport', by.y = 'airport')
```

# Map of USA and her States borders

```{r}
s <- map_data("state")
ggplot(s, aes(x = long, y= lat, group = group, fill = region)) +
  geom_polygon(colour = 'black') +
  coord_map('polyconic') +
  guides(fill = FALSE)

Num_of_flights_per_airport$state <- state.name[match(Num_of_flights_per_airport$state,state.abb)]
```

# Making 'state' to all lower case so it can be merged

```{r}
Num_of_flights_per_airport$state <- tolower(Num_of_flights_per_airport$state)
```

# Merge data frames

```{r}
flights_per_airport_data <- merge(s, Num_of_flights_per_airport,
                                  by.x = 'region',
                                  by.y = 'state')
```

# Geographic plot and plot

```{r}
map1 <- ggplot(data = flights_per_airport_data, aes( x = long.x, y = lat.x,
                                                     group = group,
                                                     fill = num_flights)) +
  geom_polygon(colour = 'black') +
  coord_map('polyconic') +
  scale_fill_gradient(low = 'white', high = 'red') +
  theme_void() +
  ggtitle('Number of flights per state 2007') +
  theme(legend.position = "none")
  
map1
```

############
# Question 4

```{r}
cascading_data <- Y2007 %>% select(TailNum, DepTime, ArrTime, Origin, Dest, ArrDelay, DepDelay)
cascading_data$datesof2007 <- datesof2007
```

# Remove rows with NA values in specific column

```{r}
cascading_data <- cascading_data %>% filter_at(vars(ArrDelay, DepDelay), all_vars(!is.na(.)))

```
# Finding the highest minutes of delay for plane

```{r}
highest_flight_delay <- cascading_data %>% select(TailNum, ArrDelay)
highest_flight_delay <- aggregate(highest_flight_delay$ArrDelay, list(highest_flight_delay$TailNum), sum)
```

# N631AS plane with highest minutes of delay
# Generating departure and arrival status

```{r}
Depstatus <- cascading_data %>% select(DepDelay)

Depstatus$depstatus <- as.factor(ifelse(Depstatus$DepDelay < 0, 'Early Departure',
                                        ifelse(Depstatus$DepDelay == 0, 'On Time', 
                                               ifelse(Depstatus$DepDelay > 0, 'Late Departure',
                                                      'nil'))))

Arrstatus <- cascading_data %>% select(ArrDelay)

Arrstatus$arrstatus <- as.factor(ifelse(Arrstatus$ArrDelay < 0, 'Early Arrival',
                                        ifelse(Arrstatus$ArrDelay == 0, 'On Time', 
                                               ifelse(Arrstatus$ArrDelay > 0, 'Late Arrival',
                                                      'nil'))))

cascading_data = subset(cascading_data, select = -c(ArrDelay, DepDelay) )

cascading_data["DepStatus"] <- Depstatus$depstatus

cascading_data["ArrStatus"] <- Arrstatus$arrstatus

```

# Filtering N631AS out from data frame

```{r}
N631ASflight <- cascading_data[cascading_data$TailNum == "N631AS",]
N631ASflight <- N631ASflight[N631ASflight$datesof2007 == "2007-01-02",] # selecting a day to focus on

print(N631ASflight)
```
############
# Question 5
# Model that predicts delays for specified day of the year
# Random forest

```{r}
library(randomForest)
library(caret)
```

# Select relevant variables from dataframe

```{r}
rfdata <- Y2007 %>% select(Month, DayofMonth, CRSDepTime,
                       CRSArrTime, DepDelay, ArrDelay, Origin,
                       Dest, Distance, CRSElapsedTime, ActualElapsedTime, TaxiIn, TaxiOut, AirTime,
                       Diverted, CarrierDelay, WeatherDelay, NASDelay, SecurityDelay, LateAircraftDelay)
```

# Remove NA values

```{r}
rfdata <- na.omit(rfdata)
```

# Delayed when > 0

```{r}
Delayed = ifelse(rfdata$ArrDelay > 0, "Yes", "No")

rfdata <- subset(rfdata, select = c(-ArrDelay))

rfdata <- data.frame(rfdata, Delayed)

rfdata$Delayed <- as.factor(rfdata$Delayed)
```

# Choosing Jan 1st as my Random Forest data

```{r}
rfdata <- filter(rfdata, Month == 1 & DayofMonth == 1)
```

# Data Partition, train/test set

```{r}
set.seed(145)

indpt_sample <- sample(2, nrow(rfdata), replace = TRUE, prob = c(0.7, 0.3))

train <- rfdata[indpt_sample == 1,]
test <- rfdata[indpt_sample == 2,]
```

# Random Forest

```{r}
set.seed(220)

rf <- randomForest(Delayed~., data = train)

rf
```
# OOB error = 12.53%

# Prediction and Confusion Matrix (train data)

```{r}
p1 <- predict(rf, train)

confusionMatrix(p1,  train$Delayed)#Accuracy : 0.999, 95% CI : (0.9996, 1)
                                   #Sensitivity : 1.0000

```

# Prediction and Confusion Matrix (test data)

```{r}
p2 <- predict(rf, test)

confusionMatrix(p2,  test$Delayed)#Accuracy : 0.876, 95% CI : (0.8672, 0.8844)
                                  #Sensitivity : 0.8959
```

# Error rate of Random Forest

```{r}
plot(rf) #200 trees
```

# Tune mtry

```{r}
t <- tuneRF(train[,-20], train[,20],
       stepFactor = 0.5,
       plot = TRUE,
       ntreetry = 200,
       trace = TRUE,
       improve = 0.05) # mtry value = 8
```

# Re-Run Tuned Random Forest

```{r}
rf <- randomForest(Delayed~., data = train,
                   ntree = 200,
                   mtry = 8,
                   importance = TRUE,
                   proximity = TRUE)

rf
```

# OOB error = 11.26% (IMPROVED)

# Number of nodes for trees

```{r}
hist(treesize(rf),
     main = "No. of Nodes for the Trees 2007",
     col = "green")                       # Distribution of nodes from 900 - 1250
                                          # Majority of the trees have an average of 80 nodes
```

# Variable Importance

```{r}
varImpPlot(rf,
           sort = T,
           n.var = 10,
           main = "Top 10 - Variable Importance 2007")
```

                                      ###END for 2007###

rm(list = ls(all.names = TRUE))       ##################
 
                                      ##START of 2006###

# Read in the data for 2006

```{r}
Y2006 <- read.csv("2006.csv")

str(Y2006)

summary(Y2006)
```

# Convert the Y2006 dates to year/month/date format 

```{r}
datesof2006 <- Y2006 %>% select(Year, Month,DayofMonth)
datesof2006 <- mutate(datesof2006,Dates=paste(Year,Month,DayofMonth,sep="/"))
datesof2006 = subset(datesof2006, select = -c(Year, Month, DayofMonth))
datesof2006 = ymd(datesof2006$Dates)
```

# Adding Y2006 dates & day of the week to flight.timing df

```{r}
flight.timingsof2006 <- Y2006 %>% select(ArrDelay)
flight.timingsof2006$datesof2006 <- datesof2006
weekday <- weekdays(datesof2006)
flight.timingsof2006$weekday <- weekday
```

# Reduce the size of the data by finding the mean of the delays per week

```{r}
weeksof2006 <- strftime( datesof2006, format = "%V")    # Converts the time to character objects
weeksof2006 <- as.numeric(weeksof2006)
flight.timingsof2006 <- cbind(flight.timingsof2006, weeksof2006)
```

# Preparation of data for loop function to find the mean delay of the 52 weeks

```{r}
WeeklyArrDelayMean <- vector(mode = "integer", length = length(1:52)) # Empty vector to store values for loop
```

# Loop for mean delay of the 52 weeks

```{r}
for (i in 1:52) {
  WeekArrDelayMeandata <- flight.timingsof2006 %>% filter(weeksof2006 == i)
  
  WeeklyArrDelayMean[i] <- mean(WeekArrDelayMeandata$ArrDelay, na.rm = TRUE)
  
} 
```

# Preparing necessary vectors for plot

```{r}
Numofweeks <- 1:52 # Generating number of weeks vector

weeklyMeanDelaysof2006 <- cbind(WeeklyArrDelayMean, Numofweeks )

weeklyMeanDelaysof2006 <- as.data.frame(weeklyMeanDelaysof2006)
```

# Plotting for mean delay of the 52 weeks

```{r}
w <- ggplot(weeklyMeanDelaysof2006,
            aes(x = Numofweeks,
                y = WeeklyArrDelayMean)
)
```

# Line plot

```{r}
w <- w + 
  geom_line()
```

# Plot Labels and plot

```{r}
w <- w + 
  labs(
    title = "Weekly Delays for 2006",
    x = "Number of weeks",
    y = "Mean of Weekly Delay"
  )

w
```

# Preparation of data for loop function to find daily mean delays

# Generating data frame

```{r}
days <- yday(datesof2006) - 1 # Generating numbered days of the year from dates

days_of_delay <- Y2006 %>% select(ArrDelay) %>% cbind(days) 

dailyArrDelayMean <- vector(mode = "integer", length = length(0:364)) # Empty vector to store values for loop
```

# Loop for daily mean of delays

```{r}
for (i in 0:364) {
  dailyArrDelayMeandata <- days_of_delay %>% filter(days == i)
  
  dailyArrDelayMean[i] <- mean(dailyArrDelayMeandata$ArrDelay, na.rm = TRUE) 
  
} # Loop for daily mean of delays
```

# Preparing necessary vectors for plot

```{r}
ymd_of_2006 <- seq(as.Date("2006-01-01"),as.Date("2006-12-31"),by="1 day") # Generating Year 2006 dates

day_of_the_week <- wday(ymd_of_2006, label = TRUE) # Days

months_of_2006 <- month(ymd_of_2006, label = TRUE) # Months
```

# Combining vectors into data frame without compromising class

```{r}
dailyArrDelayMean <- data.frame(dailyArrDelayMean = dailyArrDelayMean,
                                ymd_of_2006 = ymd_of_2006,
                                months_of_2006 = months_of_2006,
                                day_of_the_week = day_of_the_week)
```


# Plotting daily mean delays 2006

```{r}
p <- ggplot(data = dailyArrDelayMean,
            mapping = aes(x =months_of_2006,
                          y =day_of_the_week,
                          fill =dailyArrDelayMean))
```

# heatmap

```{r}
p <- p + 
  geom_tile()
```

# Colours for heatmap   

```{r}
p <- p + 
  scale_fill_distiller(limits = c(-40, 40),
                       palette = 'RdBu')
```

# Labels for heatmap and plot

```{r}
p <- p +
  labs(
    title = "Daily Delays for 2006",
    fill = "Intensity of \n mean delays",
    x = "Month",
    y = "Day"
  )

p
```

# Best time of the day
# Lowest mean delays from heat map is
# January - Wednesday

```{r}
CRSDepTime2006 <- Y2006$CRSDepTime
Month_2006 <- Y2006$Month
Delays_2006 <- Y2006$ArrDelay

Scheduled_DepTiming2006 <- data.frame(CRSDepTime2006 = CRSDepTime2006,
                                      Delays_2006 = Delays_2006,
                                      Month_2006 = Month_2006,
                                      weekday = weekday)
```

# Filter flights on October Mondays out

```{r}
JanWedScheduled_DepTiming2006 <- filter(Scheduled_DepTiming2006, Month_2006 == 1)

JanWedScheduled_DepTiming2006 <- filter(JanWedScheduled_DepTiming2006, weekday == "Wednesday")
```

# Scatter plot

```{r}
q <- ggplot(JanWedScheduled_DepTiming2006,
            aes(x = CRSDepTime2006,
                y = Delays_2006)
)
q <- q + 
  geom_point()
```

# Labels for scatter plot

```{r}
q <- q +
  labs(
    title = "January, Wednesday departures for 2006",
    x = "Time (Hrs)",
    y = "Delays (in Mins)"
  )
```

# Scale for plot and plot

```{r}
q <- q + 
  scale_x_continuous(limits = c(0,2400), breaks = 200*(0:12)) +
  scale_y_continuous(limits = c(-80,200), breaks = 20*(-10:10))

q
```

############
# Question 2

# Read in data

```{r}
planedata <- read.csv("plane-data.csv")

planedelays <- Y2006 %>% select(TailNum, ArrDelay)
```

# Remove rows with NA values in specific column

```{r}
planedelays <- planedelays %>% filter_at(vars(ArrDelay), all_vars(!is.na(.)))
```

# Group by tail number and find the mean of the delays for each tail number

```{r}
mean_plane_delays <- planedelays %>% group_by(TailNum) %>% summarise(mean_delays = mean(ArrDelay))
```

# Merging 2 data frames together by their tail number

```{r}
merged_tailnum_delays <- merge(mean_plane_delays, planedata, by.x = "TailNum", by.y = "tailnum")                                                                 

merged_tailnum_delays <- merged_tailnum_delays %>% select(TailNum, mean_delays, year)

```

# Plotting daily mean delays 2006

```{r}
r <- ggplot(data = merged_tailnum_delays,
            mapping = aes(x =mean_delays,
                          y =year
            ))
```

# Scatter plot

```{r}
r <- r + 
  geom_point()
```

# Labels for scatter plot

```{r}
r <- r +
  labs(
    title = "2006 Delays of against age of aircrafts",
    x = "Mean delays (mins)",
    y = "Year of aircraft issued"
  )
```

# Scale for plot and plot

```{r}
r <- r + 
  scale_y_discrete(limits = factor(1956:2007), breaks = 2*(978:1004)) +
  scale_x_continuous(limits = c(-50,100), breaks = 10*(-5:100))


r
```

# old planes < 1995

```{r}
OLDmerged_tailnum_delays <- merged_tailnum_delays %>% filter(year < 1995)

OLDmerged_tailnum_delays %>% summarise(mean_delays = mean(mean_delays)) #10.22
```

#new planes =>1995

```{r}
NEWmerged_tailnum_delays <- merged_tailnum_delays %>% filter(year > 1994)

NEWmerged_tailnum_delays %>% summarise(mean_delays = mean(mean_delays)) #9.66
```

############
# Question 3

```{r}
airportdata <- read.csv("airports.csv")

location_of_flights <- Y2006 %>% select(Origin, Dest, FlightNum, UniqueCarrier)

location_of_flights <- merge(airportdata, location_of_flights, by.x = 'iata', by.y = 'Origin')
```

# Number of flights per airport origin

```{r}
Num_of_flights_per_airport <- location_of_flights %>%
  group_by(airport) %>%
  summarize(num_flights = n()) 
```

# Merge data frames

```{r}
Num_of_flights_per_airport <- merge(Num_of_flights_per_airport, airportdata, by.x = 'airport', by.y = 'airport')
```

# Map of USA and her States borders

```{r}
s <- map_data("state")
ggplot(s, aes(x = long, y= lat, group = group, fill = region)) +
  geom_polygon(colour = 'black') +
  coord_map('polyconic') +
  guides(fill = FALSE)

Num_of_flights_per_airport$state <- state.name[match(Num_of_flights_per_airport$state,state.abb)]
```

# Making 'state' to all lower case so it can be merged

```{r}
Num_of_flights_per_airport$state <- tolower(Num_of_flights_per_airport$state)
```

# Merge data frames

```{r}
flights_per_airport_data <- merge(s, Num_of_flights_per_airport,
                                  by.x = 'region',
                                  by.y = 'state')
```

# Geographic plot and plot

```{r}
map1 <- ggplot(data = flights_per_airport_data, aes( x = long.x, y = lat.x,
                                                     group = group,
                                                     fill = num_flights)) +
  geom_polygon(colour = 'black') +
  coord_map('polyconic') +
  scale_fill_gradient(low = 'white', high = 'red') +
  theme_void() +
  ggtitle('Number of flights per state 2006') +
  theme(legend.position = "none")

map1
```
############
# Question 4

```{r}
cascading_data <- Y2006 %>% select(TailNum, DepTime, ArrTime, Origin, Dest, ArrDelay, DepDelay)
cascading_data$datesof2006 <- datesof2006
```

# Remove rows with NA values in specific column

```{r}
cascading_data <- cascading_data %>% filter_at(vars(ArrDelay, DepDelay), all_vars(!is.na(.)))
```

# Finding the highest minutes of delay for plane

```{r}
highest_flight_delay <- cascading_data %>% select(TailNum, ArrDelay)
highest_flight_delay <- aggregate(highest_flight_delay$ArrDelay, list(highest_flight_delay$TailNum), sum)
```

# N17175 plane with highest minutes of delay
# Generating departure and arrival status

```{r}
Depstatus <- cascading_data %>% select(DepDelay)

Depstatus$depstatus <- as.factor(ifelse(Depstatus$DepDelay < 0, 'Early Departure',
                                 ifelse(Depstatus$DepDelay == 0, 'On Time', 
                                        ifelse(Depstatus$DepDelay > 0, 'Late Departure',
                                               'nil'))))

Arrstatus <- cascading_data %>% select(ArrDelay)

Arrstatus$arrstatus <- as.factor(ifelse(Arrstatus$ArrDelay < 0, 'Early Arrival',
                                        ifelse(Arrstatus$ArrDelay == 0, 'On Time', 
                                               ifelse(Arrstatus$ArrDelay > 0, 'Late Arrival',
                                                      'nil'))))

cascading_data = subset(cascading_data, select = -c(ArrDelay, DepDelay) )

cascading_data["DepStatus"] <- Depstatus$depstatus

cascading_data["ArrStatus"] <- Arrstatus$arrstatus
```

# Filtering N17175 out from data frame

```{r}
N17175flight <- cascading_data[cascading_data$TailNum == "N17175",]
N17175flight <- N17175flight[N17175flight$datesof2006 == "2006-01-02",] # selecting a day to focus on

print(N17175flight)
```
############
# Question 5
# Model that predicts delays for specified day of the year
# Random forest

```{r}
rfdata <- Y2006 %>% select(Month, DayofMonth, CRSDepTime,
                           CRSArrTime, DepDelay, ArrDelay, Origin,
                           Dest, Distance, CRSElapsedTime, ActualElapsedTime, TaxiIn, TaxiOut, AirTime,
                           Diverted, CarrierDelay, WeatherDelay, NASDelay, SecurityDelay, LateAircraftDelay)
```

# Remove NA values

```{r}
rfdata <- na.omit(rfdata)
```

# Delayed when > 0

```{r}
Delayed = ifelse(rfdata$ArrDelay > 0, "Yes", "No")

rfdata <- subset(rfdata, select = c(-ArrDelay))

rfdata <- data.frame(rfdata, Delayed)

rfdata$Delayed <- as.factor(rfdata$Delayed)
```

# Choosing Dec 31st as my Random Forest data

```{r}
rfdata <- filter(rfdata, Month == 12 & DayofMonth == 31)
```

# Data Partition, train/test set

```{r}
set.seed(145)

indpt_sample <- sample(2, nrow(rfdata), replace = TRUE, prob = c(0.7, 0.3))

train <- rfdata[indpt_sample == 1,]
test <- rfdata[indpt_sample == 2,]
```

# Random Forest

```{r}
set.seed(220)

rf <- randomForest(Delayed~., data = train)

rf
# OOB error = 11.23%
```

# Prediction and Confusion Matrix (train data)

```{r}
p1 <- predict(rf, train)

confusionMatrix(p1,  train$Delayed)#Accuracy : 0.999, 95% CI : (0.9994, 1)
#Sensitivity : 1.0000
```

# Prediction and Confusion Matrix (test data)

```{r}
p2 <- predict(rf, test)

confusionMatrix(p2,  test$Delayed)#Accuracy : 0.883, 95% CI : (0.8737, 0.8916)
#Sensitivity : 0.9200
```

# Error rate of Random Forest

```{r}
plot(rf) #200 trees
```

# Tune mtry

```{r}
t <- tuneRF(train[,-20], train[,20],
            stepFactor = 0.5,
            plot = TRUE,
            ntreetry = 200,
            trace = TRUE,
            improve = 0.05) # mtry value = 8
```

# Re-Run Tuned Random Forest

```{r}
rf <- randomForest(Delayed~., data = train,
                   ntree = 200,
                   mtry = 8,
                   importance = TRUE,
                   proximity = TRUE)

rf
# OOB error = 10.14% (IMPROVED)
```

# Number of nodes for trees 

```{r}
hist(treesize(rf),
     main = "No. of Nodes for the Trees 2006",
     col = "green")                       # Distribution of nodes from 760 - 990
                                          # Majority of the trees have an average of 45 nodes
```

# Variable Importance

```{r}
varImpPlot(rf,
           sort = T,
           n.var = 10,
           main = "Top 10 - Variable Importance 2006")
```

###END for 2006###

rm(list = ls(all.names = TRUE))